Comparing rolling period forecasts of Facebooks' Prophet model to an ARIMA model: A Diebold-Mariano evaluation of the FTSE/JSE Top40 Index

### 2017-11-06 template/Draft_Paper.Rmd
Final draft with all edits

---
# IMPORTANT: Change settings here, but DO NOT change the spacing. 
# Remove comments and add values where applicable. 
# The descriptions below should be self-explanatory
title: "Comparing Facebook's Prophet Forecasting to an ARIMA Model: A Diebold-Mariano Evaluation of the Top40 Index"
subtitle: "A Diebold-Mariano Evaluation of the FTSE/JSE Top40 Index"
Author: "Marvelous Mubenesha"  # First Author
Student: "MBNMAR005"
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 5 # Adjust default figure sizes. This can also be done in the chunks of the text
    fig_height: 4
abstract: |
  Facebook's data science team open-sourced Prophet, a package that allows analysts to forecast a wide range of business time series at-scale. Prophet employs intensive Bayesian modelling in two tiers. Purely through MAP parameter estimation, and in a quasi-Bayesian form through an automated forecast evaluation that enables analyst to improve the model if it falls short when compared to other models. This paper assess the predictive accuracy of Prophets' forecasts to those of an ARIMA model that was identified using the Box-Jenkins methodology. This is achieved through a Diebold-Mariano evaluation of the rolling period forecast errors of returns on the FTSE/JSE Top40 Index across daily, weekly and monthly time horizons. The results of the study suggest that there is insufficient evidence to conclude that the forecasting models have unequal predictive ability within the scope of the JSE/Top40 Index. 
---
<!-- First: Set your default preferences for chunk options: -->

@@

In business, including that involving the stock market, analysts require a large number of time series forecasts to inform decision making. The challenge with existing reliable approaches to forecasting time series data are that the models require a considerable background in statistics to build and adapt to prevailing market conditions when needed. Several authors have recommended automatic forecasting, which refers to algorithms that can automatically forecast a large number of univariate time series. However, the forecasts generated by these methods have been found to be brittle as well as inflexible in allowing for prior information to be incorporated into the forecasting model [@taylor2017forecasting]. In February of 2017, Facebook's Core Datascience team open sourced ,Prophet, their time series forecasting tool whose main selling point is that it enables "forecasting-at-scale". A term coined by the data science team which is defined as, "an approach that allows a large number of analysts to forecast a large number and variety of business time series" [@taylor2017forecasting]. It has the added advantage of being less 'expensive'
than other alternatives. In addition, it is specifically prefarable over ARIMA models because of its non-linearity, flexibility and ability to accomodate varying time intervals [@taylor2017forecasting].

The objective of this paper is to compare the rolling period forecasts generated by Prophet (which uses automated, intensive Bayesian Modelling to 'forecast-at-scale'), to those obtained by an ARIMA model (constructed through the Box-Jenkins methodology),
through a Diebold-Mariano evaluation of the JSE Top40 Index. The study is particularly useful because analysts from a wide background can automatically forecast stock price returns, and incorporate their prior knowledge in the model , without a wholesome understanding of the statistical intricacies involved.

The paper is organized as follows. A review of literature relating to stock price forecasting with ARIMA models and other alternative approaches is introduced to give the reader background information on research relating to the subject matter. Thereafter, theory relating to the mathematical formulation and estimation of both the ARIMA and Prophet models is included in sections 2.2 and 2.3 respectively, to build the readers understanding of what the models look like and how they operate. Information relating to the data that was used in this study is contained in section 3, after which, the methodology used to answer the research question is described in section 4. Section 5, contains the results obtained when the methodology was executed, whilst section 6 includes a discussion of the results and how they can be interpreted to answer the research question. Lastly, the conclusion in section 7 highlights the main findings of the study and recommends further areas of research.

# Background 
In this section we will briefly review literature that relates to the comparison of different approaches to forecasting stock returns. This will include both traditional approaches as well new methods such as automatic forecasting, machine learning techniques and quasi-Bayesian forms which incorporate analysts' prior information. The review of literature is concluded with a brief introduction to the Diebold-Mariano evaluation and an outline of its suitability for the purpose of this study. The theoretical and mathematical formulation of an ARIMA forecasting model is  then briefly introduced. Thereafter, the theoretical framework and mathematical formulation of Facebooks' Prophet forecasting tool is discussed. Specific attention is drawn to Prophets distinguising features to give the reader an understanding of the value that this forecasting technique can potentially add to stock price prediction. Lastly, the Diebold-Mariano Evaluation technique is introduced as a method of formally testing superior predictive accuracy between two forecasting models.

@@

ARIMA models have been widely used as a standard model to forecast financial data and have been shown to yield acceptable forecasting errors [@adebiyi2014comparison]. @adebiyi2014comparison used the Box-Jenkins methodology to build short-term stock price prediction models for stocks on the New York Stock Exchange(NYSE) and the Nigerian Stock Exchange(NSE). Their results showed that the prediction errors of the model were within acceptable bounds.

ARIMA models rely on the assumption that residuals are homoskedastic(they have a constant variance) and normally distributed, however, some financial data displays heteroskedasticity making GARCH models more applicable. This theory is backed by a study conducted by @Muten2014 who compared the ability of ARIMA and GARCH models to forecast stock prices on the Zimbabwean Stock Exchange (ZSE). They found that the GARCH model outperforms the ARIMA model which suggests that the residuals are heteroskedastic. This is a case in point for emerging market stocks. The researchers noted that poor liquidity in the market could be a cause of the results they observed [@Muten2014].

@@

Artificial Neural Networks (ANNs) are a multi-layered perceptron as illustrated in  figure 1.  
![](ANN.png)  
*Figure 1 : Illustration of an Artificial neural network* [@tanikic2012artificial]  
An ANN consists of a sorted triple $(K, A, \omega)$. $K$ is a set representing the multiple layers/levels of the network i.e. input, hidden and output levels. A is a set of pairs that represents index for the estimate in the $i^{th}$ node of the $k^{th}$ layer of the network. In the enlarged section of Figure 1, $a_{i,k} \in A$. Lastly, $\omega$ represents a function that defines the weights ($\omega_{i,r}$) of connections between neurons/nodes [@kriesel2007brief]. Artificial Neural networks apply iterated optimization of model parameters across the network in the form of weights conditional on observed values to learn [@segaran2007programming]. Several studies have compared ARIMA model forecasts to ANN and their conclusions are contradictory.

Researchers have found that the performance of either method depends on the nature of the data and forecasting problem [@kihoro2004seasonal]. Stock price data is proposed to be nonlinear which suggests that nonlinear approaches have the potential to produce better forecasts than linear models. Furthermore, ANNs make no assumptions about the distribution of the errors as compared to the linear ARIMA model [@adebiyi2014comparison]. @adebiyi2014comparison compared NYSE stock index forecasts of an ARIMA model to those of an ANN and found that the forecasting accuracy of the ANN model was superior to that of the ARIMA model. It is evident that ANN are preferred to ARIMA models as a model free, nonlinear alternative. However, the model construction of an ANN requires trial and error to initialize parameter estimates and these parameters are not easily interpretable by analysts [@moreno2011artificial]. Researchers have found different ways to guide analysts in this respect. In 1996, @wang1996stock proposed an ARIMA-based ANN to forecast the medium-term price of the Taiwan Stock Exchange Weighted Stock Index (TSEWSI). They used the Box-Jenkins methodology to difference the series and then trained the data on a neural network with initialisations that were guided by their observations of the ARIMA model. This approach yielded forecasts with an acceptable prediction accuracy based on residual analysis of the out-of-sample data. Wang and Leu therefore concluded that the ARIMA-based Neural Network outperformed a Neural Network trained using raw stock price data [@wang1996stock]. @zhang2009stock used a combination of the backpropagation algorithm from Neural Networks with Improved Bacterial Chemotaxis Optimization (IBCO) to build a model that forecasts the S&P 500 stock index by minimizing the mean square error. Model forecasts were evaluated through simulation experiments and the results led to the conclusion that the hybrid model produced superior forecasts [@zhang2009stock]. This study further highlighted the potential that nonlinear approaches have in forecasting stock prices. An issue that arises with such methods is the complexity of the proposed models which limits the flexibility of unseasoned analysts to adjust model parameters as a way of improving forecasting accuracy. As we will see, Prophet forecasting elegantly deals with this issue in the form of a nonlinear, Generalized Additive Model (GAM).

###Quasi-Bayesian methods incorporating Analysts' knowledge

The stock price forecasting approaches considered so far purely apply technical data analysis methods to generate stock price forecasting models. However, studies that have compared pure technical approaches to those that incorporate the analysts knowledge in a quasi-Bayesian form have been shown to be superior at predicting market prices [@givoly1984quality \& @guerard1989combining]. The first instance of combining time-series model forecasts and an analysts forecasts to obtain superior forecasts of a stocks annual earnings can be dated to 1989. @guerard1989combining used an additive model to combine consensus security analyst forecasts from the S&P Annual Earnings forecaster and annual earnings forecasts generated by an ARIMA model with a constant. The combined model that was estimated using ordinary least squares reduced the mean square error of the time series and analysts forecasts from 1.28 and 1.27, respectively to 1.04. These results suggest that analysts can substantially reduce forecasting errors by combining both approaches [@guerard1989combining]. @zahedi2015application applied the same overarching principle when they used principle component analysis to determine an appropriate input variable, which they then used to train an ANN in an attempt to predict stock prices on the Tehran Stock Exchange. Their model yielded acceptable forecasting errors and performed better than the pure ANN which further reiterates the potential of incorporating an analysts knowledge to improve stock price forecasts, in both linear and nonlinear models.

## ARIMA Model Formulation

The ARIMA model is a generalisation of Autoregressive Moving Average (ARMA) models which combine autoregression and moving average features [@box1970time]. In the context of returns, the ARIMA formulation for the return at time index $t$, $r_t$, is given by,
$$  r_t = \phi_1 r_{t-1} + \phi_2 r_{t-2} + ...+ \phi_p r_{t-p} + \epsilon_t +
        \psi_1 \epsilon_{t-1} + \psi_2 \epsilon_{t-2} + ... +\psi_q \epsilon_{t-q} $$  
where:  
$\bullet \phi_i$ and $\psi_j$ are the parameters to be estimated;  
$\bullet r_{t-k}$ is the $k^{th}$ lagged return;   
$\bullet \epsilon_t$ is the error term at time $t$ which is assumed to be white noise and  
$\bullet\epsilon_{t-k}$ is the $k^{th}$ lagged error.   
$\bullet$p is the order of the autoregressive component (dependenceon history of $r_t$)
$\bullet$q is the order of the moving average component (dependenceon history of $\epsilon_t$)

@@

## Prophet Model Formulation and Estimation

The challenge with many forecasting methods such as the Box-Jenkins methodology and machine-learning approaches are that they usually require a sufficient statistical background. Furthermore, these approaches tend to be inflexible with respect to incorporating prior information. Though this can be partially achieved by initializing parameter estimates for some methods (such as ANNs) ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates [@taylor2017forecasting].  
Prophet is an **automatic forecasting** tool that uses a Bayesian Generalized Additive model to generate time series forecasts. This is in comparison to conventional methods such as the linear stochastic dependence that exists in ARIMA models. Prophet combines a configurable model that includes performance analysis evaluation with the interaction of an analyst in the loop [@taylor2017forecasting]. The model is configurable which allows an analyst to incorporate their knowledge of the behaviour of the series into the model building process through easily interpretable initial parameters that can be modified and interactive feedback when forecasts under-perform [@taylor2017forecasting].

@@

  The first step of the forecasting procedure begins with the analyst simply specifying a general model. Prophet then goes into the automated process of estimating model parameters and doing forecast evaluation. If forecasts are considered problematic based on evaluations of the Simulated Historical Forecast (SHF) errors, then the package will surface these problems. The process then transitions to the top left of the cycle in Figure 2 labelled analyst-in-the-loop. Here, the analyst can then use visually presented aspects of the results that raised the issues to modify model parameters before a better model is then estimated. The loop is repeated until forecast evaluation surfaces no problems.

![Figure 2](analyst in the loop.png){ width=100% }  
*Figure 2: Illustration of Prophets' automated forecasting procedure with an analyst in the loop* [@taylor2017forecasting]

The interpretability of Prophets' model parameters stems from the mathematical formulation of the forecasting model. The formulation posits that the series is generated by an additive, parametric function of time $r(t)$, defined by,
$$ r(t) = g(t) + s(t) +  h(t) + \epsilon_t.$$ The additive components include growth, seasonality and a component that adjusts for holidays as well as once-off events e.g. anticipated market shocks in the instance of financial time series.

The growth component, $g(t)$, is modelled using a generalized form of the logistic population growth model and is of the form;  
$$g(t) = \frac{C(t)}{1+exp(-(k+\boldsymbol{a}(t)^T \delta(t-(b+\boldsymbol{a}(t)^T \gamma)))}$$  
Where:  
$\bullet C(t)$ represents the carrying capacity of the growth component and can be modelled using a polynomial function of time, the simplest being a constant or linear model.  
$\bullet (k+\boldsymbol{a}(t)^T\delta)$ is the growth rate factor with the $\boldsymbol{a}(t)^T\delta$ term enabling the forecaster to choose where the growth rate changes( i.e change points) and   
$\bullet (b+\boldsymbol{a}(t)^T \gamma)$ is the adjusted offset parameter.

The growth component of the model has two useful features that allow the carrying capacity and growth rate of the model to vary with time. This enables an analyst to manually define when and how the growth rate changes at different change points [@taylor2017forecasting]. This feature is particularly useful for forecasting stock price data since analysts are able to incorporate market movements that lead to the growth rate either decreasing, increasing or becoming constant.  
The seasonality component, $s(t)$ with periodicity $P$ is given by,
$$s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})$$

Yearly and weekly seasonality is modelled using the standard Fourier series. Analysts can thus use prior knowledge to account for the effects of periodicities they have observed seasonally which is similar to the periodicity of Seasonal ARIMA models.

@@

Prophet translates model parameter estimation into a curve-fitting exercise by employing Maximum Apriori Posterior (MAP) estimation, a Bayesian approach to estimating optimal parameter values for the forecasting model. MAP estimation finds the maximum posterior estimates for the parameters by using the likelihood function and an apriori distribution to evaluate a posterior distribution which is then optimized to return point estimates of the parameters [@taylor2017forecasting]. Thereafter, the package automatically evaluates the forecasts by comparing the Simulated Historical Forecast (SHF) errors of baseline forecasting methods (including ARIMA forecasts) to those of Prophet. SHF errors are forecast errors generated from random points in the history of the time series model. In Prophet, SHF errors are compared visually to allow the analysts to easily adjust the interpretable parameters before automatically estimating the model and evaluating forecasts.

##Diebold-Mariano Evaluation
As is evident from sections 2.2 and 2.3, the modelling approach and assumptions of Prophet differ to those of an ARIMA model. Since the two methods differ, the Diebold-Mariano evaluation is preferred as a means of comparing the forecast accuracy of the two different methods. The Diebold-Mariano evaluation is a **model free** hypothesis test for the equal predictive accuracy of two forecasting methods. It is applicable to a wide range of situations from forecast residuals that are non-Gaussian, asymmetric loss functions, serially correlated errors and multi-period forecasts [@Mariano2000]. The Diebold-Mariano test (DM test) operates by evaluating the forecast errors for both model 1 and model 2 (i.e $e_{t_1}$ and $e_{t_2}$) at a loss function $g(\bullet)$. The loss function can be choosen flexibly as long as it is zero when the error is zero, is stricly positive for all values on the real line and, it increases with an increase in the absolute value of the error size. The loss differential between two forecasts at time $t$ is then defined by, $$ d_t =  g(e_{t1}) - g(e_{t2})$$ for $t = (1, 2, ..., M)$ where $M$ is the total number of forecasts that have been generated from each model. We then test whether, $$ \mu = E[d_t] = 0 $$ by conducting the following hypothesis test;  
Null: $$ E[d_t] = 0  \,\,\,\, \forall t$$   
Alternative : $$ E[d_t] \neq 0 $$  
Test statistic: $$DM_{stat} = \frac{\sqrt M (\bar{d} - \mu)}{\sqrt{\hat{\gamma_d(0)}}} $$

$\bar{d}$ is the sample mean of the $M$ loss differentials and $\frac{\hat{\gamma_d(0)}}{M}$ is the asymptotic variance of  ($\bar{d} - \mu$) with $\hat{\gamma_d(k)}$ representing the autocovariance of the loss differential series at lag $k=0$. Asymptotically, the $DM_{stat}$ has a standard normal distribution if the loss differential series is short memory and covariance stationary [@Mariano2000].  
There are two caveats to using the DM Test for the purposes of this paper. Firstly, the choice of the loss function depends on whether the forecast errors can be assumed to be symmetric or asymmetric. When the forecast errors are symmetric, a linear or quadratic loss function can be used to evaluate the loss differential. For forecast errors that are asymmetric, a loss fuction that satisfies the conditions outlined in the first part of this section can be used. Secondly, literature does not clearly stipulate the sample size range that is sufficient for the asymptotic distribution assumption to  hold. This implies that the DM test results must be used conservatively. <<<ASK HANJO IF HE KNOWS OF ANY PAPERS, I'VE BEEN OUT OF LUCK WITH MY SEARCH>>>

@@

plot(index, type = "l",
     main = "FTSE/JSE Top40 Index 07/2012 - 07/2017")
```
*Figure 3: Realisation of the FTSE/JSE Top40 Price Index over the period of study*

@@

table <- xtable(data, caption = "AIC for Candidate ARIMA models")
  print.xtable(table, 
             # tabular.environment = "longtable",
             floating = TRUE,
             table.placement = 'H', 
             # scalebox = 0.3, 
             comment = FALSE,
             include.rownames = getOption("xtable.include.rownames", FALSE), #exclude row number
             caption.placement = 'bottom'
             )

>>>>

--- 
# IMPORTANT: Change settings here, but DO NOT change the spacing. 
# Remove comments and add values where applicable. 
# The descriptions below should be self-explanatory
title: "Comparing rolling period forecasts of Facebooks' Prophet model to an ARIMA model: A Diebold-Mariano evaluation of the FTSE/JSE Top40 Index"
subtitle: "A Diebold-Mariano Evaluation of the FTSE/JSE Top40 Index"
Author: "Marvelous Mubenesha"  # First Author
Student: "MBNMAR005"
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 5 # Adjust default figure sizes. This can also be done in the chunks of the text
    fig_height: 4
abstract: |
  Facebook's data science team open-sourced Prophet, a package that allows analysts to forecast a wide range of business time series at-scale. Prophet employs intensive Bayesian modelling in two tiers. Purely through MAP parameter estimation, and in a quasi-Bayesian form through an automated forecast evaluation that enables analyst to improve a forecasting model if it underperforms when compared to other traditional models. This paper assesses the predictive accuracy of Prophets' forecasts to those of an ARIMA model that was identified using the Box-Jenkins methodology. This is achieved through a Diebold-Mariano evaluation of the rolling period forecast errors of returns on the FTSE/JSE Top40 Index across daily, weekly and monthly time horizons. The results of the study suggest that there is insufficient evidence to conclude that the forecasting models have unequal predictive ability within the scope of the FTSE/JSE Top40 Index. 
---
<!-- First: Set your default preferences for chunk options: -->

@@

In business, including that involving the stock market, analysts require a large number of time series forecasts to inform decision making. The challenge with existing reliable approaches to forecasting time series data are that the models require a considerable background in statistics to build and adapt to prevailing market conditions when needed. Several authors have recommended automatic forecasting, which refers to algorithms that can automatically forecast a large number of univariate time series. However, the forecasts generated by these methods have been found to be brittle as well as inflexible in allowing for prior information to be incorporated into the forecasting model [@taylor2017forecasting]. In February of 2017, Facebook's Core Datascience team open sourced Prophet, a time series forecasting tool with unique features that enable "forecasting-at-scale". Forecasting-at-scale is a term coined by the data science team which they define as, "an approach that allows a large number of analysts to forecast a large number and variety of business time series" [@taylor2017forecasting]. It has the added advantage of being less 'expensive'
than other alternatives. In addition, it is specifically preferable over ARIMA models because of its non-linearity, flexibility and ability to accomodate varying time intervals [@taylor2017forecasting].

The objective of this paper is to compare the rolling period forecasts generated by Prophet (which uses automated, intensive Bayesian Modelling to 'forecast-at-scale'), to those obtained from an ARIMA model (constructed through the Box-Jenkins methodology),
through a Diebold-Mariano evaluation of the FTSE/JSE Top40 Index. The study is particularly useful because analysts from a wide background can reliably forecast stock price returns automatically, and incorporate their prior knowledge in the forecasting model, without a wholesome understanding of the statistical intricacies involved. It is hoped that this could create opportunities for automated invesment mechanisms.

The paper is organized as follows. A review of literature relating to stock price forecasting with ARIMA models and other alternative approaches is introduced in section 2.1 to give the reader background information on existing research relating to the subject matter. Thereafter, theory relating to the mathematical formulation and estimation of an ARIMA and Prophet model is included in sections 2.2 and 2.3 respectively. Subsequently, in section 2.4, the theoretical motivation and mathematical formulation of the Diebold-Mariano test is introduced as a means of comparing the predictive accuracy of two prediction models. Information relating to the data that was used in this study is contained in section 3, after which, the methodology used to answer the research question is described in section 4. Section 5, contains the results obtained when the methodology was executed, whilst section 6 includes a discussion of the results. The discussion includes a response to the research question as well as some useful insights from the results. Furthermore, it highlights some recommendations for further areas of research based on the findings of the paper. Lastly, the conclusion in section 7 summarizes the results of the study relative to the main objective of the paper.

# Background 
In this section we will briefly review literature that relates to the comparison of different approaches to forecasting stock returns. This will include both traditional approaches as well as relatively new methods such as automatic forecasting, machine learning techniques and quasi-Bayesian forms which incorporate analysts' prior information. The theoretical and mathematical formulation of an ARIMA forecasting model is  then briefly introduced. Thereafter, the theoretical framework and mathematical formulation of Facebooks' Prophet forecasting tool is discussed. Specific attention is drawn to Prophets distinguising features to give the reader an understanding of the value that this forecasting technique can potentially add to stock price prediction. The background information is concluded with an introduction to the Diebold-Mariano Evaluation as a suitable framework to formally compare the predictive accuracy of two forecasting models.

@@

ARIMA models have been widely used as a standard model to forecast financial data and have been shown to yield forecasting errors that lie within acceptable bounds [@adebiyi2014comparison]. @adebiyi2014comparison used the Box-Jenkins methodology to build short-term stock price prediction models for stocks on the New York Stock Exchange(NYSE) and the Nigerian Stock Exchange(NSE). Their results showed that the predictive power of the ARIMA model was satisfactory based on forecast error diagnostics.

ARIMA models rely on the assumption that residuals are homoskedastic(they have a constant variance) and normally distributed. However, the residual errors of some financial data display heteroskedasticity making GARCH models more applicable. This theory is backed by a study conducted by @Muten2014 who compared the ability of ARIMA and GARCH models to forecast stock prices on the Zimbabwean Stock Exchange (ZSE). They found that the GARCH model outperforms the ARIMA model which suggests that incorporating heteroskedasticity of the residual errors improves the forecasting ability of the model. This is a case in point for emerging market stocks. The researchers noted that poor liquidity in the market could be a cause of the results they observed [@Muten2014].

@@

An Artificial Neural Networks (ANNs) is a multi-layered perceptron with nodes that simulate the action of neurons as illustrated in  figure 1.  
![](ANN.png)  
**Figure 1 : Illustration of an Artificial neural network** [@tanikic2012artificial]  
Formally, an ANN consists of a sorted triple $(K, A, \omega)$. $K$ is a set representing the multiple levels of the network i.e. input, hidden and output levels. A is a set of pairs that represents the index for the estimated parameter at the $i^{th}$ node of the $k^{th}$ level in the network. This is illustrated by the curved down arrow in Figure 1 which points to an enlarged segment of the $i^{th}$ node in the $k^{th}$ level i.e. $a_{i,k} \in A$. Lastly, $\omega$ represents a function that defines the weights ($\omega_{i,r}$) of connections between nodes that interact across adjacent levels [@kriesel2007brief]. Artificial Neural networks apply iterated optimization of model parameters across the network in the form of weights conditional on observed values to learn [@segaran2007programming]. Several studies have compared ARIMA model forecasts to ANN and their conclusions are contradictory.

Researchers have found that the performance of either method depends on the nature of the data and forecasting problem [@kihoro2004seasonal]. Stock price data is proposed to be nonlinear which suggests that nonlinear approaches have the potential to produce better forecasts than linear models. Furthermore, ANNs make no assumptions about the distribution of the errors as compared to the linear ARIMA model [@adebiyi2014comparison]. @adebiyi2014comparison compared NYSE stock index forecasts of an ARIMA model to those of an ANN and found that the forecasting accuracy of the ANN model was superior to that of the ARIMA model. It is evident that ANN are preferred to ARIMA models as a model free, nonlinear alternative. However, the model construction of an ANN requires trial and error to initialize parameter estimates and these parameters are not easily interpretable by analysts [@moreno2011artificial]. Researchers have found different ways to guide analysts in this respect. In 1996, @wang1996stock proposed an ARIMA-based ANN to forecast the medium-term price of the Taiwan Stock Exchange Weighted Stock Index (TSEWSI). They used the Box-Jenkins methodology to difference the series and then trained the data on a neural network with initialisations that were guided by their observations of the ARIMA model. Their hybrid model yielded forecasts that were comparably superior to the individual models used in isolation based on an analysis of the out-of-sample residual error diagnostics. @wang1996stock therefore concluded that the ARIMA-based Neural Network outperformed a Neural Network trained using raw stock price data [@wang1996stock]. @zhang2009stock used a combination of the backpropagation algorithm from Neural Networks with Improved Bacterial Chemotaxis Optimization (IBCO) to build a model that forecasts the S&P 500 stock index by minimizing the mean square error. Model forecasts were evaluated through simulation experiments and the results led to the conclusion that the hybrid model produced superior forecasts [@zhang2009stock]. This study further highlighted the potential that nonlinear approaches have in forecasting stock prices. An issue that arises with such methods is the complexity of the proposed models which limits the flexibility of unseasoned analysts to adjust model parameters as a way of improving forecasting accuracy. As we will see, Prophet elegantly deals with this issue in the form of a nonlinear, Generalized Additive Model (GAM).

###Quasi-Bayesian forecasting methods that incorporate an analysts' knowledge

The stock price forecasting approaches considered so far purely apply technical data analysis methods to generate stock price forecasting models. However, studies that have compared pure technical approaches to those that incorporate an analysts knowledge in a quasi-Bayesian form have been shown to be superior at predicting market prices [@givoly1984quality \& @guerard1989combining]. The first instance of combining time-series model forecasts and an analysts forecasts to obtain superior forecasts of a stocks annual earnings can be dated to 1989. @guerard1989combining used an additive model to combine consensus security analyst forecasts from the S&P Annual Earnings forecaster and annual earnings forecasts generated by an ARIMA model with a nonzero mean. The combined model that was estimated using ordinary least squares reduced the mean square error of the time series and analysts forecasts from 1.28 and 1.27, respectively to 1.04. These results suggest that analysts can reduce forecasting errors by combining technical methods with expertise developed through market knowledge [@guerard1989combining]. @zahedi2015application applied the same overarching principle when they used principle component analysis to determine an appropriate input variable which they then used to train an ANN in an attempt to predict stock prices on the Tehran Stock Exchange. Their model yielded reduced forecasting errors and performed better than the pure ANN. These results further reiterate the potential of incorporating an analysts knowledge to improve stock price forecasts, in both linear and nonlinear models.

## ARIMA model formulation

The ARIMA model is a generalisation of Autoregressive Moving Average (ARMA) models which combine autoregression and moving average features [@box1970time]. In the context of returns, the ARIMA formulation for the return at time index $t$, $r_t$, is given by,
$$  r_t = \phi_1 r_{t-1} + \phi_2 r_{t-2} + ...+ \phi_p r_{t-p} + \epsilon_t +
        \psi_1 \epsilon_{t-1} + \psi_2 \epsilon_{t-2} + ... +\psi_q \epsilon_{t-q} $$where:  
$\bullet$  $\phi_i$ and $\psi_j$ are the parameters to be estimated  
$\bullet$  $r_{t-k}$ is the $k^{th}$ lagged return   
$\bullet$  $\epsilon_t$ is the error term at time $t$ which is assumed to be white noise  
$\bullet$  $\epsilon_{t-k}$ is the $k^{th}$ lagged error   
$\bullet$  $p$ is the order of the autoregressive component (dependence on history of $r_t$) and  
$\bullet$  $q$ is the order of the moving average component (dependence on history of $\epsilon_t$).

@@

## Prophet model formulation and estimation

The challenge with many forecasting methods such as the Box-Jenkins methodology and machine learning approaches discussed in the literature review are that they usually require a sufficient statistical background to construct and modify if necessary. Furthermore, some of these approaches tend to be inflexible with respect to incorporating prior information. In some instances, such as with ANN, incorporating prior information can be partially achieved through initializing parameter estimates. However, an analyst with limited statistical training can have difficulty interpreting and thus modifying initial parameter estimates [@taylor2017forecasting].  
Prophet is an **automatic forecasting** tool that uses a Bayesian Generalized Additive model to generate time series forecasts. This is in comparison to conventional methods such as the linear stochastic dependence that exists in ARIMA models. Prophet combines a configurable model that includes an automated evaluation of forecasts with the interaction of an analyst in the loop [@taylor2017forecasting]. The model is configurable which allows an analyst to incorporate their knowledge of the behaviour of the series into the model building process through easily interpretable initial parameters that can be modified and interactive feedback when forecasts under-perform [@taylor2017forecasting].

@@

  The first step of the forecasting procedure begins with the analyst simply specifying a general model (i.e. modelling). Prophet then goes into the automated process of estimating model parameters and performing forecast evaluation.  Automated forecast evaluation is undertaken by constructing baseline forecasts using methods such as the sample mean, ARIMA, exponential smoothing and naive estimates (e.g. last value or seasonal last value). Thereafter, simulated historical forecasting (SHF) errors are evaluated by comparing forecasts of the baseline methods and prophet to the observed historical values. SHF errors are forecast errors generated by random points in the history of a time series model. Prophets SHF errors that are relatively large compared to the baseline methods are then surfaced with visual illustrations of the results. The process then transitions from surface problems to the top left of the cycle in Figure 2 labelled analyst-in-the-loop. Here, the analyst can then visually inspect forecasts to identify issues and modify features of the Prophet model before it is re-estimated in the modelling stage. The loop is repeated until forecast evaluation surfaces no problems [@taylor2017forecasting].

![Figure 2](analyst in the loop.png){ width=80% }  
**Figure 2: Illustration of Prophets' automated forecasting procedure with an analyst in the loop** [@taylor2017forecasting]

The interpretability of Prophets' model parameters stems from the mathematical formulation of the generalised additive model. The formulation posits that the series is generated by an additive, parametric function of time $r(t)$, defined by,
$$ r(t) = g(t) + s(t) +  h(t) + \epsilon_t.$$ The additive components include growth, seasonality and a component that adjusts for holidays as well as once-off events e.g. anticipated market shocks in the instance of financial time series.

Firstly, the growth component, $g(t)$, is modelled using a generalized form of the logistic population growth model and is of the form;  
$$g(t) = \frac{C(t)}{1+exp(-(k+\boldsymbol{a}(t)^T \delta(t-(b+\boldsymbol{a}(t)^T \gamma)))}$$  
where:  
$\bullet$   $C(t)$ represents the carrying capacity of the growth component and can be modelled using a polynomial function of time, the simplest being a constant or linear model.  
$\bullet$   $(k+\boldsymbol{a}(t)^T\delta)$ is the growth rate factor with the $\boldsymbol{a}(t)^T\delta$ term enabling the forecaster to choose where the growth rate changes( i.e change points) and   
$\bullet$   $(b+\boldsymbol{a}(t)^T \gamma)$ is the adjusted offset parameter.  
The growth component of the model has two useful features that allow the carrying capacity and growth rate of the model to vary with time. This enables an analyst to manually define when and how the growth rate changes at different change points [@taylor2017forecasting]. This feature is particularly useful for forecasting stock price data since analysts are able to incorporate market movements that lead to the growth rate either decreasing, increasing or becoming constant.

Secondly, the seasonality component, $s(t)$ with periodicity $P$ is given by,
$$s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})$$

Yearly and weekly seasonality is modelled using the standard Fourier series. Analysts can thus use prior knowledge to account for the effects of periodicities they have observed seasonally.

@@

Prophet translates model parameter estimation into a curve-fitting exercise by employing Maximum Apriori Posterior (MAP) estimation, a Bayesian approach to estimating optimal parameter values for the forecasting model. MAP estimation finds the maximum posterior estimates for the parameters by using the likelihood function and an apriori distribution to evaluate a posterior distribution which is then optimized to return point estimates of the parameters [@taylor2017forecasting].

##Diebold-Mariano evaluation
As is evident from sections 2.2 and 2.3, the modelling approach and assumptions of Prophet differ to those of an ARIMA model. Since the two methods differ, the Diebold-Mariano evaluation is preferred as a means of formally comparing the forecasting accuracy of the two models statistically. The Diebold-Mariano evaluation is a **model free** hypothesis test for the equal predictive accuracy of two forecasting methods, referred to as model 1 and model 2. It is applicable to a wide range of situations from forecast residuals that are non-Gaussian, asymmetric loss functions, serially correlated errors and multi-period forecasts [@Mariano2000]. The Diebold-Mariano test (DM test) operates by evaluating the forecast errors for both model 1 and model 2 (i.e $e_{t_1}$ and $e_{t_2}$) at a loss function $g(\bullet)$. The loss function can be choosen flexibly as long as it is zero when the error is zero, is strictly positive for all values on the real line and, it increases with an increase in the absolute value of the error size. The loss differential between two forecasts at time $t$ is then defined by, $$ d_t =  g(e_{t1}) - g(e_{t2})$$ for $t = (1, 2, ..., M)$ where $M$ is the total number of forecasts that have been generated from each model. We then test whether, $$ \mu = E[d_t] = 0 $$ by conducting the following hypothesis test;  
Null hypothesis: $$ E[d_t] = 0  \,\,\,\, \forall t$$   
Alternative hypothesis: $$ E[d_t] \neq 0 $$  
Test statistic: $$DM_{stat} = \frac{\sqrt M (\bar{d} - \mu)}{\sqrt{\hat{\gamma_d(0)}}} $$

$\bar{d}$ is the sample mean of the $M$ loss differentials and $\frac{\hat{\gamma_d(0)}}{M}$ is the asymptotic variance of  ($\bar{d} - \mu$) with $\hat{\gamma_d(k)}$ representing the autocovariance of the loss differential series at lag $k=0$. Asymptotically, the $DM_{stat}$ has a standard normal distribution if the loss differential series is short memory and covariance stationary [@Mariano2000].  
There are two caveats to using the DM Test for the purposes of this paper. Firstly, the choice of the loss function depends on whether the forecast errors can be assumed to be symmetric or asymmetric. When the forecast errors are symmetric, a linear or quadratic loss function can be used to evaluate the loss differential. For forecast errors that are asymmetric, a loss fuction that satisfies the conditions outlined in the first part of this section can be used. Secondly, literature does not clearly stipulate the sample size range that is sufficient for the asymptotic distribution assumption to  hold. In light of this, relatively large sample sizes were used in this study and the DM test results are applied conservatively.

@@

plot(index, type = "l",
     main = "FTSE/JSE Top40 Index 07/2012 - 07/2017")
```
**Figure 3: Realisation of the FTSE/JSE Top40 Price Index over the period of study**

@@

table <- xtable(data)
  print.xtable(table, 
             # tabular.environment = "longtable",
             floating = TRUE,
             table.placement = 'H', 
             # scalebox = 0.3, 
             comment = FALSE,
             include.rownames = getOption("xtable.include.rownames", FALSE) #exclude row number
             )

```
## Forecast error evaluation
